---
title: "In-class Exercise 4"
description: |
  A short description of the post.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Installing and loading required packages
- sf is mainly used for importing geospatial data
- raster is for handling 

```{r echo=TRUE, eval=TRUE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap', 'tidyverse')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
# onemapsgapi
```

## 1.1 More information about the packages used
- **maptools**: Manipulate geographic data. 
  - Will be used to convert spatial objects into ppp format of spatstat
  
- **sf**: used for importing, managing, and processing **geospatial** data
  - specifically **vector-based** geospatial data
  
- **raster**: reads, writes, manipulates, analyses and model of **gridded** spatial data

- **spatstat**: contains useful functions for point pattern analysis. 
   - Will be used to perform **1st- and 2nd-order spatial point patterns analysis** and derive kernel density estimation (**KDE**) layer.
   
- **tmap**: provides functions for plotting cartographic **quality** *static* point patterns maps or *interactive* maps by using leaflet API.

- **tidyverse** (for CHILDCARE dataset): used for importing, wrangling and visualising data and consists of a family of packages:
  - **readr** for importing csv data,
  - **readxl** for importing Excel worksheet,
  - **tidyr** for manipulating data,
  - **dplyr** for transforming data, and
  - **ggplot2** for visualising data

# 2. Import Geospatial Data
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
sg_sf <- st_read(dsn="data/shapefile", layer="CostalOutline") 
mpsz_sf <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

# 3. Importing Aspatial Data
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

## 3.1 Converting from aspatial to geospatial
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

```

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}

```
