---
title: "In-class Exercise 3"
description: |
  In this in-class exercise, I learn about analytical mapping. 
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Installing and loading required packages
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# 2. Import Data
```{r echo=TRUE, eval=TRUE}
mpszpop2020 <- read_rds("data/rds/mpszpop2020.rds")
```


# 3. Plot a quantile map of initial data

- Can see that there are missing values in AGED

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, fig.retina=3, fig.height=4}
tm_shape(mpszpop2020)+
  tm_fill("AGED",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```
# 4. Visualise Extreme Values 

## 4.1 Box plot
 
```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE, fig.retina=3, fig.align='center', fig.height=5, fig.width=4}
ggplot(data=mpszpop2020, 
       aes(x = "", 
           y = AGED)) +
  geom_boxplot()

```


## 4.2 Percentile Map
- Usually 6 categories
- Extreme values are much better highlighted
 
### 4.2.1 Extract a variable from sf data frame

#### 4.2.1.1 Exclude records with NA
```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  drop_na()
```


#### 4.2.1.2 See range of DEPENDENCY (*Extra: For my understanding*)
```{r echo=TRUE, eval=TRUE}
range(mpszpop2020a$DEPENDENCY)
```


#### 4.2.1.3 See quartiles of DEPENDENCY (*Extra: For my understanding*)
```{r echo=TRUE, eval=TRUE}
quantile(mpszpop2020a$DEPENDENCY)
```

#### 4.2.1.4 Defining a different set of quartiles
```{r echo=TRUE, eval=TRUE}
percent <- c(0, .01, .1, .5, .9, .99, 1)
var <- mpszpop2020a["DEPENDENCY"] %>% 
  st_set_geometry(NULL)
quantile(var[,1], percent)
```


### 4.2.2 Creating the get.var function
```{r echo=TRUE, eval=TRUE}
get.var<-function(vname, df) {
  v <- df[vname]%>%
  st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```


### 4.2.3 Plotting a percentile map using tmap functions

- This plot has empty holes (because we removed NA)
- This is what we normally do

```{r echo=TRUE, eval=TRUE}
percent <- c(0, .01, .1, .5, .9, .99, 1)
var <- get.var("DEPENDENCY", mpszpop2020a)
bperc <- quantile(var, percent)
tm_shape(mpszpop2020a) + 
  tm_polygons() + 
  tm_shape(mpszpop2020a) + 
  tm_fill("DEPENDENCY", 
          breaks=bperc, 
          palette="Blues",
          labels=c("< 1%", "1%-10%", "10%-50%", "50%-90%", "90%-99%", ">99%"))+
  tm_borders() +
  tm_layout(title = "Percentile Map", title.position = c("right", "bottom"))

```


### 4.2.4 Plotting a percentile map using a function

- This plot will not have empty holes (since we have filled it)
- The idea is to Plot as a polygon first then the fill function so that won't get awkward spaces after removing NA values

```{r echo=TRUE, eval=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01, .1,.5,.9,.99,1)
  var <- get.var(vnam,df)
  bperc <- quantile (var, percent)
  tm_shape (mpszpop2020) + 
    tm_polygons() + 
    tm_shape(df)+
    tm_fill(vnam, title=legtitle, breaks=bperc, palette="Blues", 
            labels = c("< 1%", "1%-10%", "10%-50%", "50%-90%", "90%-99%", ">99%"))+
     tm_borders() +
    tm_layout(title = mtitle, title.position=c("right", "bottom"))
    
}
```


#### 4.2.4.1 Test the percentile mapping function - AGED
```{r echo=TRUE, eval=TRUE}
percentmap("AGED", mpszpop2020a)
```

#### 4.2.4.2 Test the percentile mapping function - YOUNG
```{r echo=TRUE, eval=TRUE}
percentmap("YOUNG", mpszpop2020a)
```

## Just for my understanding (to see the difference in map after choosing a different palette): *Please ignore this section*
### A percentile mapping function (PRGn palette)

```{r echo=TRUE, eval=TRUE}

## Plot as a polygon first 
# then fill function so that you wont get the awkward spaces after removing NA values
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01, .1,.5,.9,.99,1)
  var <- get.var(vnam,df)
  bperc <- quantile (var, percent)
  tm_shape (mpszpop2020) + 
    tm_polygons() + 
    tm_shape(df)+
    tm_fill(vnam, title=legtitle, breaks=bperc, palette="PRGn", 
            labels = c("< 1%", "1%-10%", "10%-50%", "50%-90%", "90%-99%", ">99%"))+
     tm_borders() +
    tm_layout(title = mtitle, title.position=c("right", "bottom"))
    
}
```


### Test function
```{r echo=TRUE, eval=TRUE}
percentmap("AGED", mpszpop2020a)
```


## 4.3  Box Map 
- Helps to identify the lower and upper outliers

### 4.3.1 Creating the boxbreaks function

```{r echo=TRUE, eval=TRUE}
  # identify outliers using upper and lower percentile
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```


### 4.3.2 Test the boxbreak function
```{r echo=TRUE, eval=FALSE}
var <- get.var("AGED", mpszpop2020a) 
boxbreaks(var)
```

### 4.3.3 Ensure that AGED doesn't contain 0 (although already removed previously)
```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  filter(AGED>=0)
var <- get.var("AGED", mpszpop2020a)
boxbreaks(var)
```

### 4.3.4 Creating *Boxmap* function (This will have holes)
```{r echo=TRUE, eval=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

### 4.3.5 Plot box map - ECONOMY ACTIVE (Have holes)
```{r fig.retina=4, fig.height=5}
boxmap("ECONOMY ACTIVE", mpszpop2020a)
```


### 4.3.6 Tidy version of box map (No holes, nicer to see)
```{r echo=FALSE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(mpszpop2020) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

### 4.3.7 Plot box map - ECONOMY ACTIVE (Have holes)
```{r fig.retina=4, fig.width=12, fig.height=6}
boxmap("ECONOMY ACTIVE", mpszpop2020a)
```

## 4.4 Choropleth Map for Rates

### 4.4.1 Comparing distribution of total and aged population, 2020

### 4.4.1.1 TOTAL
```{r echo=FALSE, eval=TRUE, fig.retina=4, fig.height=5}
qtm(mpszpop2020, "TOTAL",
    title = "Distribution of total population, 2020")
```

### 4.4.1.2 AGED
```{r echo=FALSE, eval=TRUE,fig.retina=4, fig.height=5}
qtm(mpszpop2020, "AGED",
    title = "Distribution of aged population, 2020")
```


## 4.5 Raw Rate map
```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>%
  mutate(`AGED%` = (`AGED`/`TOTAL`) * 100) %>%
  filter (`AGED%` >= 0)

```

## 4.5.1 Plot Raw Rate map
```{r echo=FALSE, eval=TRUE, message=FALSE, fig.retina=4, fig.height=5}
var <- get.var("AGED%", mpszpop2020a)
boxbreaks(var)
boxmap("AGED%",mpszpop2020a)
```


## 4.6 Compare Absolute and Rate Choropleth Maps

### 4.6.1 Absolute
```{r echo=FALSE, fig.retina=4, fig.height=5}
boxmap("AGED",mpszpop2020a)
```

### 4.6.2 Rate
```{r echo=FALSE, fig.retina=4, fig.height=5}
boxmap("AGED%",mpszpop2020a)
```

* There is a difference in the colours.

