---
title: "Hands-on Exercise 2"
description: |
  In this hands-on exercise, I learn how to handle geospatial data in R by using sf package and performing data science tasks using tidyverse package.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 08-26-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Getting Started  


This code chunk performs 3 tasks:
- A list called packages will be created and will consists of all the R packages required to accomplish this hands-on exercise.
- Check if R packages on package have been installed in R and if not, they will be installed.
- After all the R packages have been installed, they will be loaded.  


```{r echo=TRUE, eval=TRUE}
packages = c('sf', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Importing Geospatial Data

## Importing **polygon** feature data in shapefile format
```{r echo=TRUE, eval=TRUE}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

## Importing **polyline** feature data in shapefile form
```{r echo=TRUE, eval=TRUE}
cyclingpath = st_read(dsn = "data/geospatial", layer = "CyclingPath")

```

## Importing GIS data in kml format
```{r echo=TRUE, eval=TRUE}
preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```

# Checking the Content of A Simple Feature Data Frame

## st_geometry() function
```{r echo=TRUE, eval=TRUE}
st_geometry(mpsz)
```

## glimpse() function
```{r echo=TRUE, eval=TRUE}
glimpse(mpsz)
```

## head() function
```{r echo=TRUE, eval=TRUE}
head(mpsz, n=5)  
```

# Plotting the Geospatial Data

## plot() function
```{r echo=TRUE, eval=TRUE}
plot(st_geometry(mpsz))
```

## Choose plot by a specific attribute
```{r echo=TRUE, eval=TRUE}
plot(mpsz["PLN_AREA_N"])
```


# Working with Projection

## Assigning EPSG code to a simple feature data frame

### st_crs function - Retrieve coordinate reference system from object
```{r echo=TRUE, eval=TRUE}

st_crs(mpsz)

```

### Assign correct EPSG code
```{r echo=TRUE, eval=TRUE}
mpsz3414 <- st_set_crs(mpsz, 3414)
```

### Check CSR again
```{r echo=TRUE, eval=TRUE}
st_crs(mpsz3414)
```

## Transform the original data from geographic coordinate system to projected coordinate system.
```{r echo=TRUE, eval=TRUE}
preschool3414 <- st_transform(preschool, crs = 3414)
```


# Importing and Converting An Aspatial Data

```{r echo=TRUE, eval=TRUE}
listings <- read_csv("data/aspatial/listings.csv")

```

```{r echo=TRUE, eval=TRUE}
glimpse(listings) 
```

## Creating a simple feature data frame from an aspatial data frame
```{r echo=TRUE, eval=TRUE}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>% st_transform(crs = 3414)
```


```{r echo=TRUE, eval=TRUE}
glimpse(listings_sf) 
```


# Geoprocessing with sf package

```{r echo=TRUE, eval=TRUE}
buffer_cycling <- st_buffer(cyclingpath, dist=5, nQuadSegs = 30)
```


```{r echo=TRUE, eval=TRUE}
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)

```

# Point-in-polygon count

## Identify pre-schools located inside each Planning Subzone - st_intersects(). 

## Calculate numbers of pre-schools that fall inside each planning subzone using length() of Base R  
```{r echo=TRUE, eval=TRUE}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
```

## List planning subzone with most number of pre-school
```{r echo=TRUE, eval=TRUE}
top_n(mpsz3414, 1, `PreSch Count`)
```

## Calculate the density of pre-school by planning subzone.
```{r echo=TRUE, eval=TRUE}
mpsz3414$Area <- mpsz3414 %>% st_area()
```


```{r echo=TRUE, eval=TRUE}
mpsz3414 <- mpsz3414 %>% mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# Explorotary Data Analysis (EDA)

## Histogram
```{r echo=TRUE, eval=TRUE}
hist(mpsz3414$`PreSch Density`)
```

## Histogram using ggplot
```{r echo=TRUE, eval=TRUE}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```

Done! :)

