---
title: "Hands-on Exercise 3"
description: |
  In this hands-on exercise, I learn how to plot functional and truthful choropleth maps by using an R package called tmap package.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 08-29-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Importing Packages
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```


## 2. Importing data in R

### 2.1 Importing Geospatial Data into R

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.   

```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```


### 2.2 Examine the content of `mpsz`


```{r echo=TRUE, eval=TRUE}
mpsz
```


### 2.3 Import Attribute Data into R


```{r, echo=TRUE, message=FALSE}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## 2.4 Data Preparation
Prepare data table with variables like:
-PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY
-(PA and SZ fields can be used as unique identifiers to geocode to MP14_SUBZONE_WEB_PL shapefile)*

### 2.4.1 Data Wrangling

```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```


### 2.5 Join the attribute data and geospatial data

- Must convert to uppercase since one column is in uppercase and the other is not. 
```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```


### 2.6 Join the geographical data and attribute table


```{r echo=TRUE, eval=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```



```{r}
mpszpop2020 <- write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```



## 3.0 Choropleth Mapping Geospatial Data Using tmap

### 3.1 Draw a cartographic standard choropleth map as shown below.


```{r echo=TRUE, eval=TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```


### 3.2 Create choropleth map by using *tmap*'s elements


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```


### 3.3 Drawing a base map


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```


### 3.4 Drawing a choropleth map using *tm_polygons()*


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```


### 3.5 Drawing a choropleth map using *tm_fill()* and *tm_border()**


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```


### 3.6 Add the boundary of the planning subzones


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```


### 3.7 Plotting choropleth maps with built-in classification methods


### 3.7.1 **Quantile** data classification


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```


###  3.7.2 **Equal** data classification


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```


###  3.7.3 PLOTTING CHOROPLETH MAP WITH CUSTOM BREAK

### 3.7.4 Show descriptive statistics of **DEPENDENCY** field


```{r echo=TRUE, eval=TRUE}
summary(mpsz_pop2020$DEPENDENCY)
```


### 3.7.5 Plot the choropleth map


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```


# 4.0 Changing Colour Scheme

# 4.1 Change the colour


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```


# 4.2 Reverse the colour shading


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```


# 4.2 Map layouts

# 4.2.1 Map legend

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```


# 4.2.2 Classic map style


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```


# 4.2.3 Draw other map furniture (Cartographic)
 

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```


# 4.2.4 Reset the default style


```{r echo=TRUE, eval=TRUE}
tmap_style("white")
```


## 4.3 Drawing Small Multiple Choropleth Maps 

## 4.3.1 Small multiple choropleth maps are created defining ***ncols*** in **tm_fill()**


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```


## 4.3.2 Small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```


## 4.3.3 Multiple small choropleth maps are created by using **tm_facets()**. 


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```


## 4.3.4 Multiple small choropleth maps are created by creating multiple stand-alone maps with **tmap_arrange()**. 


```{r echo=TRUE, eval=TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")
agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")
tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```


## 4.4 Mapping Spatial Object Meeting a Selection Criterion

### 4.4.1 Use selection funtion to map spatial objects meeting the selection criterion. 


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```




