---
title: "Hands-on Exercise 3"
description: |
  In this hands-on exercise, I learn how to plot functional and truthful choropleth maps by using an R package called tmap package.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 08-29-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Importing Packages
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

More about packages used:

- **sf**: used for importing, managing, and processing geospatial data
- **tidyverse**: used for importing, wrangling and visualising data. It consists of a family of R packages, such as:
  - **readr** for importing csv data,
  - **readxl** for importing Excel worksheet,
  - **tidyr** for manipulating data,
  - **dplyr** for transforming data, and
  - **ggplot2** for visualising data
- **tmap**: Used to plot high cartographic quality plots (Since plot() is mean for plotting the geospatial object for quick look)

## 2. Importing data in R
Data used:

- MP14_SUBZONE_WEB_PL, a **polygon** feature layer in ESRI shapefile format,
- **Aspatial** data file: Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format

### 2.1 Importing **Geospatial** Data into R
The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.   

```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

- From the results above we can see that:
  - there are **323 multipolygon features and 15 fields** in the sf data frame. 
  - **svy21** is the projected coordinates system
  
### 2.2 Examine the content of `mpsz`
```{r echo=TRUE, eval=TRUE}
mpsz
```


### 2.3 Import **Attribute** Data into R
```{r, echo=TRUE, message=FALSE}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

## 2.4 Data Preparation

Prepare data table with variables like:

- **YOUNG**: age group 0 to 4 until age groyup 20 to 24,
- **ECONOMY ACTIVE**: age group 25-29 until age group 60-64,
- **AGED**: age group 65 and above,
- **TOTAL**: all age group, and   
- **DEPENDENCY**: the ratio between young and aged against economy active group
- PA and SZ fields can be used as **unique identifiers** to geocode to MP14_SUBZONE_WEB_PL shapefile*

### 2.4.1 Data Wrangling
```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])+rowSums(.[12])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+rowSums(.[13:15]))%>%
  mutate(`AGED`=rowSums(.[16:21])) %>%
  mutate(`TOTAL`=rowSums(.[3:21])) %>%  
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`) /`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)
```

### 2.4.2 Data Wrangling (For my own understanding)

#### **Filter and group_by**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) 
popdata2020v2
```

#### **summarise**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  summarise(`POP` = sum(`Pop`))
popdata2020v2
```

#### **ungroup**
- Can see that there is no longer this result # Groups:   PA, SZ [332]

```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  ungroup()
popdata2020v2
```

#### **pivot_wider**
- Pivot table from skinny and long to wide and short table 

```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  pivot_wider(names_from=AG, 
              values_from=POP)
glimpse(popdata2020v2)
```

#### **mutate - YOUNG**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  mutate(YOUNG = rowSums(.[3:6])+rowSums(.[12]))
glimpse(popdata2020v2)
```

#### **mutate - ECONOMY ACTIVE**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+rowSums(.[13:15]))
glimpse(popdata2020v2)
```

#### **mutate - AGED**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  mutate(`AGED`=rowSums(.[16:21]))
glimpse(popdata2020v2)
```

#### **mutate - TOTAL**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  mutate(`TOTAL`=rowSums(.[3:21]))
glimpse(popdata2020v2)
```

#### **mutate - DEPENDENCY**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`) /`ECONOMY ACTIVE`)
glimpse(popdata2020v2)
```

#### **select()**
```{r echo=TRUE, eval=TRUE}
popdata2020v2 <- popdata2020v2 %>%
  select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`)
glimpse(popdata2020v2)
```

## 2.5 Join the attribute data and geospatial data
### 2.5.1 Convert to uppercase
- Must convert to uppercase since one column is in uppercase and the other is not. 

```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

### 2.5.2  **Join** the geographical data and attribute table using **left_join()**
```{r echo=TRUE, eval=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

### 2.5.3 Export an entire R object (metadata and all) using **write_rds()**
- To be used in ICE 3. 

```{r echo=TRUE, eval=TRUE}
mpszpop2020 <- write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## 3.0 Choropleth Mapping Geospatial Data Using tmap
### 3.1 Draw a cartographic standard choropleth map as shown below.

My own notes:

- A **choropleth map** brings together 2 datasets: **spatial** data representing a partition of geographic space into distinct districts, and **statistical** data representing a variable aggregated within each district
- qtm: means quick thematic map -> Draw a thematic map quickly
- Disadvantage of **qtm()**: it makes aesthetics of individual layers harder to control.

```{r echo=TRUE, eval=TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### 3.2 Create choropleth map by using *tmap*'s elements
- we use tmap’s drawing elements to overcome cons of qtm by drawing a high quality cartographic choropleth map

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

### 3.3 Drawing a base map
- **basic building block** of tmap is tm_shape() followed by one or more layer elemments such as tm_fill() and tm_polygons()
- this means that you need to always have + after tm_shape()

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

### 3.4 Drawing a choropleth map using **tm_polygons()**
- To draw a choropleth map showing geographical distribution of a **SELECTED VARIABLE** by planning subzone, we just need to assign the target variable 
- For e.g. **assign Dependency to tm_polygons()**
- NOTE:
  - Default colour scheme used is **YlOrRd** of ColorBrewer
  - By default, **Missing** value will be shaded in **grey.**
  
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

## 3.5 Drawing a choropleth map using **tm_fill()** and **tm_border()**
Short overview of **tm_polygons()**, **tm_fill()**, **tm_borders()**:

- **tm_polygons()** is a **wraper** of tm_fill() and tm_border(). 
- **tm_fill()** shades the **polygons** by using the default colour scheme 
- **tm_borders()** adds the **borders** of the shapefile onto the choropleth map.

### 3.5.1 Draw a choropleth map by using **tm_fill() alone**
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

### 3.5.2 Add the boundary of the planning subzones using **tm_borders()** & the previous tm_filll()
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

### 3.6 Plotting choropleth maps with built-in classification methods
- tmap provides a total 10 data classification methods, namely: fixed, sd, equal, pretty (default), quantile, kmeans, hclust, bclust, fisher, and jenks.

### 3.6.1 **Quantile** data classification that used 5 classes
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5)
```

### 3.6.2 **Equal** data classification
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

### 3.6.3 Prepare choropleth maps by using different classification methods supported by tmap and compare their differences
1. fixed, 
2. sd, 
3. equal, 
4. pretty (default), 
5. quantile, 
6. kmeans, 
7. hclust, 
8. bclust, 
9. fisher, and 
10. jenks

#### fixed
Only **fixed** have error

```{r echo=TRUE, eval=TRUE}
# tm_shape(mpsz_pop2020)+
#   tm_fill("DEPENDENCY",
#           n = 5,
#           style = "fixed") +
#   tm_borders(alpha = 0.5) +
#   tm_layout(legend.position = c("right", "top"), 
#             title= '1. fixed', 
#             title.position = c('right', 'top'))
```

#### sd
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "sd") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '2. sd', 
            title.position = c('right', 'top'))
```

#### equal
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '3. equal', 
            title.position = c('right', 'top'))
```

#### pretty
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "pretty") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '4. pretty', 
            title.position = c('right', 'top'))
```

#### quantile
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '5. quantile', 
            title.position = c('right', 'top'))
```

#### kmeans
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '6. kmeans', 
            title.position = c('right', 'top'))
```

#### hclust
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "hclust") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '7. hclust', 
            title.position = c('right', 'top'))
```

#### bclust
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "bclust") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '8. bclust', 
            title.position = c('right', 'top'))
```

#### fisher
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "fisher") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '9. fisher', 
            title.position = c('right', 'top'))
```

#### jenks
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '10. jenks', 
            title.position = c('right', 'top'))
```

### 3.6.4 **Quantile** data classification that different classes (2, 6, 10, 20)
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '2 classes', 
            title.position = c('right', 'top'))
```

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '6 classes', 
            title.position = c('right', 'top'))
```

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 10,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '10 classes', 
            title.position = c('right', 'top'))
```

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 20,
          style = "quantile") +
  tm_borders(alpha = 0.5) +
  tm_layout(legend.position = c("right", "top"), 
            title= '20 classes', 
            title.position = c('right', 'top'))
```

### 3.6.5 PLOTTING CHOROPLETH MAP WITH CUSTOM BREAK
- custom break means you define your own intervals 

#### 3.6.5.1 Show descriptive statistics of **DEPENDENCY** field
```{r echo=TRUE, eval=TRUE}
summary(mpsz_pop2020$DEPENDENCY)
```

#### 3.6.5.2 Plot the choropleth map
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

# 4.0 Changing Colour Scheme
# 4.1 Change the colour
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

# 4.2 Reverse the colour shading
- Original colour shading (light to dark green)

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

- Reversed colour shading (dark to light green)

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

# 4.3 Map layouts
- Map layout refers to: combination of all map elements into a cohensive map. 
- Map elements include: 
  - objects to be mapped, 
  - title, 
  - scale bar, 
  - compass, margins and aspects ratios.
- Colour settings and data classification methods covered in the previous section relate to the palette and break-points are used to affect how the map looks.

# 4.3.1 Map **legend**
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

# 4.3.2 Classic map **style**
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

# 4.3.3 Draw other map furniture (Cartographic)
- Other map furnitures include: **compass, scale bar and grid lines**
- tm_compass(), tm_scale_bar() and tm_grid() are used to add compass, scale bar and grid lines onto the choropleth map

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

# 4.3.4 Reset the default style
```{r echo=TRUE, eval=TRUE}
tmap_style("white")
```

## 4.4 Drawing Small Multiple Choropleth Maps 
## 4.4.1 Small multiple choropleth maps are created defining ***ncols*** in **tm_fill()**
Code below uses: tm_shape, tm_fill, tm_layout, tm_borders, tmap_style

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

## 4.4.2 Small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments
Code below uses: tm_shape, tm_fill, tm_layout

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

## 4.4.3 Multiple small choropleth maps are created by using **tm_facets()**. 
- **tm_facets**: specifies how small multiples are placed in a **grid**
Code below uses: tm_shape, tm_fill, tm_facets, tm_layout, tm_borders

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

## 4.4.4 Multiple small choropleth maps are created by creating multiple **stand-alone maps** with **tmap_arrange()**. 
- here you can see that we create multiple **stand-alone maps** FIRST
- before we proceed with tmap_arrange

```{r echo=TRUE, eval=TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")
agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")
tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

## 4.5 Mapping Spatial Object Meeting a **Selection Criterion**
### 4.5.1 Use selection funtion to map spatial objects meeting the **Selection Criterion**
- Here, the selection criterion is **REGION_N=="CENTRAL REGION"**

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

