---
title: "Hands-on Exercise 11"
description: |
  In this hands-on exercise, I learn how to model geographical accessibility using R Packages.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 10-31-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 4
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# 1. Learning Outcomes

In this hands-on exercise, I learn how to:

- to import GIS polygon data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
- to import aspatial data into R and save them as simple feature data frame by using appropriate functions of sf package of R;
- to computer accessibility measure by using Hansen’s potential model and Spatial Accessibility Measure (SAM); and
- to visualise the accessibility measures by using tmap and ggplot2 packages.

# 2. Datasets

4 data sets will be used in this hands-on exercise, they are:

- `MP14_SUBZONE_NO_SEA_PL`: URA Master Plan 2014 subzone boundary GIS data. This data set is downloaded from data.gov.sg.
- `hexagons`: A 250m radius hexagons GIS data. This data set was created by using st_make_grid() of sf package. It is in ESRI shapefile format.
- `ELDERCARE`: GIS data showing location of eldercare service. This data is downloaded from data.gov.sg. There are two versions. One in ESRI shapefile format. The other one in Google kml file format. For the purpose of this hands-on exercise, ESRI shapefile format is provided.
- `OD_Matrix`: a distance matrix in csv format. There are six fields in the data file. They are:
  - origin_id: the unique id values of the origin (i.e. fid of hexagon data set.),
  - destination_id: the unique id values of the destination (i.e. fid of ELDERCARE data set.),
  - entry_cost: the perpendicular distance between the origins and the nearest road),
  - network_cost: the actual network distance from the origin and destination,
  - exit_cost: the perpendicular distance between the destination and the nearest road), and
  - total_cost: the summation of entry_cost, network_cost and exit_cost.
  
All the values of the cost related fields are in **metres**.

# 3. Install and Launch R Packages

The R packages need for this exercise are as follows:

- Spatial data handling: sf
- Modelling geographical accessibility: spatialAcc
- Attribute data handling: tidyverse, especially readr and dplyr
- thematic mapping: tmap
- Statistical graphic: ggplot2
- Statistical analysis: ggstatsplot

```{r echo=TRUE, eval=TRUE}
packages = c('tmap', 'SpatialAcc','sf', 'ggstatsplot', 'reshape2','tidyverse')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

- Notice that with tidyverse, we do not have to install readr, dplyr and ggplots packages separately. 
- In fact, tidyverse also installs other R packages such as tidyr, stringr, forcats, tibble, purrr and magrittr.

# 4. Geospatial Data Wrangling

## 4.1 Importing geospatial data

- *st_read()* of **sf** package to import these three data sets 

```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

## 4.2 Updating CRS information
### 4.2.1 Assign correct code

- *st_transform()* updates the newly imported mpsz with the correct ESPG code (i.e. 3414)

```{r echo=TRUE, eval=TRUE}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

### 4.2.2 Verify CRS

- *st_crs()* of **sf**  package to verify the projection of the newly transformed mpsz_svy21

```{r echo=TRUE, eval=TRUE}
st_crs(mpsz)
```

Notice that the EPSG is indicated as 3414 now.

## 4.3 Clean and update attribute fields of geospatial data

- The code chunks below will be used to exclude redundant fields.
- *mutate()*  of **dplyr** package to create new field called `demand` and `capacity` that will be added into the data table of `hexagons` and `eldercare` sf data frame respectively. 

```{r echo=TRUE, eval=TRUE}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

# 5. Aspatial Data Handling and Wrangling
## 5.1 Importing Distance Matrix

- *read_cvs()* of **readr** package to import OD_Matrix.csv.
- The imported object is a tibble data.frame called `ODMatrix`.

```{r echo=TRUE, eval=TRUE}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

- The imported ODMatrix organised the distance matrix columnwise.
- On the other hand, most of the modelling packages in R is expecting where rows represent origins (i.e. AKA **from field**) and the columns represent destination (i.e. AKA **to field**)

## 5.2 Tidy distance matrix
### 5.2.1 Transform matrix

- *spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.
- **Note**: We should use *pivot_wider()* instead of *spread()*

```{r echo=TRUE, eval=TRUE}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

### 5.2.2 Convert from m to km

- The distance is measured in metre because SVY21 projected coordinate system is used. 
- The code chunk below will be used to convert the unit f measurement from metre to kilometre using *as.matrix()* of **Base R** 

```{r echo=TRUE, eval=TRUE}
distmat_km<-as.matrix(distmat/1000)
```

# 6. Modelling and Visualising Accessibility using Hansen Method
## 6.1 Computing Hansen’s accessibility

- *ac()* of **SpatialAcc** package to compute Hansen’s accessibility
- *data.frame()* is used to save the output in a data frame called acc_Handsen.

```{r echo=TRUE, eval=TRUE}
acc_Hansen <- data.frame(ac(hexcen$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

## 6.2 Rename column name

- *colnames()* of **Base R** to rename column names

```{r echo=TRUE, eval=TRUE}
colnames(acc_Hansen) <- "accHansen"
```

## 6.3 Convert to tibble format

- *tbl_df()* of **dplyr** package to convert data table into tibble format

```{r echo=TRUE, eval=TRUE}
acc_Hansen <- tbl_df(acc_Hansen)
```

## 6.4 Join *acc_Hansen* with *hexagons* 

- *bind_cols()* of **dplyr** will be used to join the acc_Hansen tibble data frame with the hexagons simple feature data frame.

```{r echo=TRUE, eval=TRUE}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## 6.5 Single code chunk version of above codes
```{r echo=TRUE, eval=FALSE}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "Hansen"))
colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- tbl_df(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

## 6.6 Visualising Hansen’s accessibility
### 6.6.1 Extract map extend

- *st_bbox()* of **sf** package to extract the extend of hexagons simple feature data frame

```{r echo=TRUE, eval=TRUE}
mapex <- st_bbox(hexagons) #view extent
```

### 6.6.2 Plot eldercare

- Uses a collection of mapping functions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## 6.7 Statistical graphic visualisation
In this section, we are going to compare the distribution of Hansen’s accessibility values by URA Planning Region.

### 6.7.1 Add the planning region field

- *st_join()* of **sf** package to add the planning region field into *haxegon_Hansen* simple feature data frame

```{r echo=TRUE, eval=TRUE}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

### 6.7.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# 7. Modelling and Visualising Accessibility using KD2SFCA Method

## 7.1 Compute KD2SFCA’s accessibility

- The code chunk below calculates Hansen’s accessibility using *ac()* of **SpatialAcc** and data.frame() is used to save the output in a data frame called `acc_KD2SFCA`.
- Notice that **KD2SFCA** is used for **family** argument.

```{r echo=TRUE, eval=TRUE}
acc_KD2SFCA <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            d0 = 50,
                            power = 2, 
                            family = "KD2SFCA"))
colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- tbl_df(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## 7.2 Visualise KD2SFCA’s accessibility

- The code chunk below uses a collection of mapping fucntions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.
- Notice that **mapex** is reused for **bbox** argument.

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```


## 7.2 Statistical graphic visualisation
Here, we will compare the distribution of KD2CFA accessibility values by URA Planning Region.

### 7.2.1 Add the planning region field

- *st_join()* of **sf** package to add the planning region field into *hexagon_KD2SFCA* simple feature data frame

```{r echo=TRUE, eval=TRUE}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

### 7.2.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# 8. Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method
## 8.1 Computing SAM accessibility

- *ac()* of **SpatialAcc** and *data.frame()* is used to save the output in a data frame called `acc_SAM`
- Notice that **SAM** is used for **family** argument.

```{r echo=TRUE, eval=TRUE}
acc_SAM <- data.frame(ac(hexagons$demand,
                         eldercare$capacity,
                         distmat_km, 
                         d0 = 50,
                         power = 2, 
                         family = "SAM"))
colnames(acc_SAM) <- "accSAM"
acc_SAM <- tbl_df(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## 8.2 Visualise SAM’s accessibility

- The code chunk below uses a collection of mapping functions of tmap package to create a high cartographic quality accessibility to eldercare centre in Singapore.
- Notice that **mapex** is reused for **bbox** argument.

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## 8.3 Statistical graphic visualisation
Now, we are going to compare the distribution of SAM accessibility values by URA Planning Region.

### 8.3.1 Add the planning region field

- *st_join()* of **sf** package to add the planning region field into *hexagon_SAM * simple feature data frame

```{r echo=TRUE, eval=TRUE}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

### 8.3.2 Plot distribution

- *ggplot()* of **ggplot2** package to plot distribution 

```{r echo=TRUE, eval=TRUE, fig.width=8, fig.height=8}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

