---
title: "In-class Exercise 5"
description: |
  A short description of the post.
author:
  - name: Nor Aisyah
    url: https://www.linkedin.com/in/nor-aisyah/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    toc_depth: 3
---


```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = FALSE)
```

To show your figure images in a higher resolution, fix the retina in the above code chunk above.

# 1. Install and load R packages
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
packages = c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
    }
  library(p,character.only = T)
}
```

More on the packages used:

- **Maptools** is for creating spatial* objects
- **sf** is for importing geospatial data
- **raster** is for creating a raster object
- **spatstat** is for performing SPPA
- **tmap** is for plotting thematic maps
- **tidyverse** is for data handling

# 2. Data used:

- **MP14_SUBZONE_WEB_PL**: a polygon feature data in ESRI shapefile format from data.gov.sg, providing information of URA 2014 Master Plan Planning Subzone boundary data
- **CHILDCARE**: a point feature data in rds format that provides both location and attribute information of childcare centres from data.gov.sg
- **CHAS**: a point feature data in rds format that provides location and attribute information of CHAS clinics


# 3. Import the Geospatial data

- Importing shapefile using **st_read()** of **sf** package. The output object is in tibble sf object class.

```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
mpsz_sf <- st_read(dsn = "data/shapefile", layer = "MP14_SUBZONE_WEB_PL")
```

We can see from the results above that mpsz_sf projection is in SVY21. 

# 4. Import aspatial data from rds folder

- **read_rds()** of **readr** package is used instead of **readRDS()** of base R is used. This is because the output of **read_rds()** is in tibble object.

```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
CHAS <- read_rds("data/rds/CHAS.rds")
childcare <- read_rds("data/rds/childcare.rds")
```

Note: There are some data issue in the childcare data frame because `Lat` and `Lng` should be in numeric data type. The coordinate fields are in decimal degrees. Hence, wgs referencing system is assumed. 

# 5. Convert aspatial data frame into sf objects

## 5.1 CHAS conversion to sf object
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
CHAS_sf <- st_as_sf(CHAS,
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs=3414)
```

Note: st_as_sf accepts the coordinates in character data type.  

## 5.2 Childcare conversion to sf object
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
childcare_sf <- st_as_sf(CHAS,
                    coords = c("Lng",
                               "Lat"),
                    crs=4326) %>%
  st_transform(crs=3414)
```

Note: In childcare we have Lng and Lat as the coordinates where they are both in demical degrees. The coordinate system for that is WGS84, crs code 4326. We then use **st_transform** to change the coordinates of a geometry from one spatial reference system to another. 

# 6. Plot to review

Notes for arguments: 
- alpha is to set transparency
- col is to set the colour of the dots
- size is to set the size of the dots

```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
tmap_mode('view')
tm_shape(childcare_sf) +
  tm_dots(alpha=0.4,
          col="blue",
          size=0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha=0.4,
          col="red",
          size=0.05)

tmap_mode('plot')
```

Note:

- Darker blue points show that you have more than 1 childcare centers
- With the interactive mode, you can:
  - Zoom in and out to examine the data 
  - Switch views to see the context like which neighbourhood they are in etc.
    - GrayCanvas to see the distribution
    - OpenStreetMap to see the local context

# 7. Geospatial Data Wrangling

## 7.1 Converting from sf to Spatial* dataframe 

- Here we are using using **as_Spatial** of sf package 
- The asterisk in Spatial* is to say that we have more than 1 Spatial objects

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

Notes: 

- The Spatial objects will consist of data table which doesn't contain the geometry object
- bbox is showing the extent of coordinates
- proj4string contains the projection of the map data
- The difference between mpsz and the other 2 childcare and CHAS is that only mpsz has the additional polygons data since the other 2 are point feature data.


## 7.2 Converting from Spatial* dataframe into Spatial* objects (sp format)

- Here we are using as.SpatialPoint() of as.SpatialPolygon() of maptools package
- maptools is only capable of converting from one Spatial object to another Spatial object. That is why we performed the above step in 7.1. 

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS,"SpatialPoints")
mpsz_sp <- as(mpsz,"SpatialPolygons")
```

Note: After performing the above, the difference that we can see in childcare and childcare_sp is that the dataframe disappears. Similarly for CHAS and mpsz.

## 7.3 Converting from Spatial* objects into spatstat ppp format

Here we are using as.ppp() of maptools package.

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```
    
Note: 

- Difference between childcare_sp and childcare_ppp is that the projections and bbox are no longer there. 
- There is also a data called window here which we can use to see our study area. Later on, we can define the study area for our owin object.

## 7.4 Remove duplicates points using **jitter** of spatstat package

- In hands on exercise 4, there were 3 ways to remove duplicates. Jitter is just one of the ways. 

### 7.4.1 Remove duplicates points in childcare
```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE,
                             nsim=1, 
                             drop = TRUE)
any(duplicated(childcare_ppp_jit))
```

The above results shows that there are no duplicates. 

### 7.4.2 Remove duplicates points in CHAS
```{r}
CHAS_ppp_jit <- rjitter(CHAS_ppp, 
                             retry=TRUE,
                             nsim=1, 
                             drop = TRUE)
any(duplicated(CHAS_ppp_jit))
```

Similarly, for CHAS, the above results shows that there are no duplicates. 

## 7.5 Extract Punggol Planning Area 
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

Note:

- When working with polygon dataframe, you can slowly type out the code above to see the fields, pausing at **$** sign.
- **IMPORTANT**: The full syntax should have the **,** at the back! If you don't have **,** it will not work.
- From the above results, we can see that we only have one object class and all of it is belonging to PUNGGOl area.

## 7.6 Convert SpatialPolygonDataFrame into SpatialPolygons object
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
pg_sp <- as(pg, "SpatialPolygons")
```

## 7.7 Convert SpatialPolygons into owin object
```{r  eval=TRUE, echo=TRUE, message=FALSE, error=FALSE}
pg_owin <- as(pg_sp, "owin")
```






